Vagueness, Noise, and Signaling

### 2015-10-02 paper/version_04/paper.tex
section 3 edits
Section 3 edits and appendix

Many concepts and expressions are vague. A vague category knows clear
cases that fall under it, clear cases that do not, and also so-called
borderline cases. Borderline cases do not clearly apply, nor do they
clearly not apply, and there may be differences between borderline
cases in terms of how well they represent the category in
question. Vagueness does not seem to dramatically affect the success
of everyday communication, but it is troublesome for some of the most
prominent theories of language and meaning. This is especially so for the logico-positivist tradition of Frege, Russell and the young
Wittgenstein which is challenged by the paradoxes vagueness gives rise
to.

@@

The next section introduces the background against which the work presented here can be
appreciated. Section~\ref{sec:repl-diff-dynam} introduces a generalization of the replicator
dynamic which is derived from the idea that agents imitate other agents' behavior while
possibly confusing similar states. This imprecise imitation dynamic is explored in
Section~\ref{sec:exploring-rdd}. Section~\ref{sec:discussion} reflects and compares our
approach to others. Appendix~\ref{sec:formal-stuff} provides formal detail.

@@

Formally, a sim-max game consists of a set of states $\States$, a set of messages $\Messgs$ typically with much
fewer messages than states, a probability distribution $\Pr \in \Delta(\States)$ such that
$\Pr(\state)$ gives the probability that state $\state$ occurs, a similarity metric\footnote{A
  metric is a function of distance between any two points in a given space. It should satisfy
  certain axioms that ensure behavior that one would intuitively expect from the words
  ``similarity'' and ``distance'' alone, but these details do not matter for the purposes of
  this paper.}
  % : similarity should always be non-negative, zero only for the distance
  % between an element and itself, symmetric, and satisfy the triangle inequality that
  % $\similarity(x,z) \le \similarity(x,y) + \similarity(y,z)$
on states $\similarity \mycolon \States \times \States \rightarrow \mathds{R}$ such that
$\similarity(\mystate{1}, \mystate{2})$ is the (perceptual) similarity between $\mystate{1}$
and $\mystate{2}$, and a utility function $\Utils \mycolon \States \times \States \rightarrow
\mathds{R}$ such that $\Utils(\mystate{1}, \mystate{2})$ is the payoff for sender and receiver
for a play with actual state $\mystate{1}$ and receiver interpretation $\mystate{2}$. We
identify the receiver's acts with the states of the world, so that the game is one of guessing
the actual state, so to speak. The utility function should be a monotonically increasing
function of similarity.
A sender's behavior can be represented by a pure strategy $\Spure \in \Messgs^\States$, which deterministically defines which message would be used for each state, a mixed strategy $\Smixed \in \Delta(\Messgs^\States)$, which characterizes a probability distribution over pure strategies, or a behavioral strategy $\Sstrat \in \Delta(\Messgs)^\States$, which defines the probability that a given message would be used for each state.
The analogous happens for the receiver's behavior.
A play in such a game with behavioral strategies is illustrated in Figure~\ref{fig:sim-max-illustration}.

@@

  \caption{A round of play in a sim-max game with behavioral strategies: Nature (N) chooses a state $\mystate{1}$ with probability $\Pr(\mystate{1})$; a sender (S) selects a message $\messg$ with probability $\Sstrat(\messg \probbar \mystate{1})$; a receiver (R) selects a state $\mystate{2}$ with probability $\Rstrat(\mystate{2} \probbar \messg)$. Payoff for both sender and receiver is given by $\Utils(\mystate{1}, \mystate{2})$.}
   %which is typically proportional to state similarity $\similarity(\mystate{1}, \mystate{2})$.}
  \label{fig:sim-max-illustration}
\end{figure}

@@

Seen in this light, sim-max games may provide a foundation to approaches in cognitive semantics
that rely on the notion of conceptual spaces.
\citet[][70--77]{Gardenfors2000:Conceptual-Spac}, for example, has prominently argued that
natural categories are convex regions in conceptual space. If the conceptual space has a
suitable metric, convex categories can be derived from a set of prototypes. The category
corresponding to prototype $p$ is the set of points that are more similar to $p$ than to any
other. In this way, \citeauthor{Gardenfors2000:Conceptual-Spac} argues, an efficient categorization
system can be obtained: storing the prototypes lets us recover the categories without having to
store each category's extension. However, what is left unexplained is where the prototypes
come from, and why we would not see just any distribution of prototypes as an equally efficient
classification system. This is where sim-max games can contribute a principled approach to
deriving, in an independent way, not only convex categories but also prototypical exemplars
belonging to them.  These ideas, and more, are developed further by
\citet{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag,OConnor2014-OCOEPC},
among others.

@@

Alternative approaches are taken by, for example, \citet{FrankeJager2010:Vagueness-Signa} and
\citet{OConnor2013:The-Evolution-o}, who show different ways how the above desiderata can be met by evolving
strategies in sim-max games. The approach we take here is similar in spirit, but different in
relevant detail. A more in-depth comparison is deferred until Section~\ref{sec:discussion}. Let
us first look into our own proposal in more detail.

@@

Signaling agents can adapt their dispositions to act, given some feedback about their past
success, in multiple ways. Usually, we would assume that changes in behavior should, at least
on average, lean towards increasing chances of communicative success. Such behavioral
adaptations can be described at different levels of abstraction. At the level of individual
agents, we can picture a more or less idealized process of how each agent adapts dispositions
for future actions based on various pieces of information available to the agent. More
abstractly, at the level of a population of agents, we can describe how average behavioral
dispositions will evolve. The population-level perspective abstracts over small stochastic
fluctuations and zooms in on the general tendency or direction of evolution that ensues from
behavior at the agent-level.

@@

The \emph{expected utility} of choices at each choice point is:\footnote{The notation
  $\propto$, for ``proportional to'', that is used here and hereafter means that the right-hand
  side might still need to be normalized so as to have probabilities sum to one. In general,
  writing $P(x) \propto f(x)$ for any function $f \mycolon X \rightarrow \mathds{R}$ is
  shorthand for $P(x) = \frac{f(x)}{\sum_{x' \in X} f(x')}$. }
\begin{align*}
  \EU(\messg, \state, \Rstrat) & = \sum_{\state \in \States}
  \Rstrat(\state' \probbar \messg) \myts U(\state, \state') \\
  \EU(\state', \messg, \Sstrat) & = \sum_{\state \in
    \States} P(\state \probbar \messg)  \myts
  U(\state, \state')\,, \text{\ \  where } &
P(\state \probbar \messg) & \propto \Pr(\state) \myts \Sstrat(\messg \probbar \state) \,.
\end{align*}
Expected utilities are sums, over all possible concrete outcomes, of the utilities of these
outcomes, weighted by how likely they occur. For instance, the expected utility for the sender
of choosing message $\messg$, given state $\state$ and a receiver strategy $\Rstrat$ is the
sum, for each interpretation $\state'$, of the probability that some random receiver will
choose $\state'$ given $\messg$ times the actual utility of triple $\state$, $\messg$ and
$\state'$.

@@

The replicator dynamic is an abstract population-level dynamic that describes the mean expected
change of behavioral dispositions in a population of signalers. There are several ways of
deriving the replicator dynamic from agent-level processes of behavioral adaptation. We focus
here on one of the simplest: \emph{imitation of success}
\citep[see][]{Sandholm2010:Population-Game}. The intuitive idea is the following (see
Appendix~\ref{sec:formal-stuff} for a derivation of the standard replicator dynamic from this update scheme).
Every now and then a random agent gets a chance to alter his behavior for one of his choice
points. Say, this is a sender who gets to ``reconsider'' his choice of message for state
$\state$. The revising agent then observes what some random agent does at $\state$, say
$\messg$, and will henceforth play $\messg$ in $\state$ with a probability given by the
expected utility of $\messg$ for state $\state$.

@@

The presence of observation and realization noise also affects imitation of successes. Here, we
focus on the main ideas. Formal detail is provided by Appendix~\ref{sec:formal-stuff}. Take a sender who gets to
revise behavior for state $\state$. What agents can plausibly revise by imitation is their pure
strategy, which maps perceived states onto messages. We should therefore assume that when an
agent gets to revise his strategy for the perceived state $\state$, this need not necessarily
be the actual state. Moreover, when that agent observes $\state$, another possibly
to-be-imitated agent may yet perceive a completely different state altogether. Given what that
latter agent perceives, his actual (pure) strategy will determine what he plays. In sum, to
describe how likely the potential imitator observes a message choice $\messg$, we are
interested in the conditional probability $P_o(\messg \probbar \state)$ that some other random
agent realizes $\messg$ when the first agent perceived $\state$. This
$P_o(\messg \probbar \state)$ is derived from the prior probability of states, the current
sender population behavior $\Sstrat$ and the given observation noise (see
Appendix~\ref{sec:formal-stuff}). Eventually, the imitating agent adopts $\messg$ as his choice for $\state$ with
probability given by the expected utility of sending $\messg$ when perceiving state
$\state$. Expected utility should, of course, take the probabilistic confusability of states
into account as well.

@@

Appendix~\ref{sec:formal-stuff} shows how imprecise imitation of this sort leads to mean changes in population
frequencies of choices that can be covered by the following discrete-time formulation:
\begin{align}
  \label{eq:nRD}
  \Sstrat'(\messg \probbar \state) & \propto P_o(\messg \probbar \state) \myts \EU(\messg,
  \state, \Rstrat)\,, & \Rstrat'(\state \probbar \messg) & \propto P_o(\state \probbar \messg)
  \myts \EU(\state, \messg, \Sstrat)\,.
\end{align}
\noindent This looks very much like the discrete-time formulation of the standard replicator
dynamic in (\ref{eq:RDDiscrete}), but there are, of course, the aforementioned
differences. Firstly, expected utility here takes stochastic confusability of states into
account. Secondly, where the standard replicator dynamic had probabilities
$\Sstrat(\messg \probbar \state)$ and $\Rstrat(\state \probbar \messg)$, we now have
$P_o(\messg \probbar \state)$ and $P_o(\state \probbar \messg)$ respectively. If there is no
observation or realization noise, $P_o(\messg \probbar \state)$ reduces to
$\sigma(\messg \probbar \state)$ and $P_o(\state \probbar \messg)$ reduces to
$\Rstrat(\state \probbar \messg)$. The noisy replicator dynamic in (\ref{eq:nRD})
conservatively extends the classic case in (\ref{eq:RDDiscrete}).

@@

Let the state space consist of $n \ge 2$ states that are equally spaced across the unit
interval, including $0$ and $1$. All states occur, for simplicity, with the same probability (\emph{i.e.}~$P_a$ is uniform).
The distance $\card{\mystate{i} - \mystate{j}}$ is the objective, physical similarity between
two states $\mystate{i}$ and $\mystate{j}$. Distance in physical space feeds into a perceptual
similarity function, as described by \citet{Nosofsky1986:Attention-Simil}:
\begin{align*}
  \similarity(\mystate{i}, \mystate{j} \myts ; \myts \imprecision) =
      \begin{cases}
    1 & \textrm{if } \imprecision = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
    0 & \textrm{if } \imprecision = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
 \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^2}{ \imprecision^2} \right ) & \textrm{otherwise,} \\
    \end{cases}
\end{align*}
where $\imprecision \ge 0$ is an imprecision parameter. When $\imprecision=0$ agents perfectly
discriminate between states; when $\imprecision \rightarrow \infty$ agents cannot discriminate
states at all. Figure~\ref{fig:NosofskySim} gives an impression of Nosofsky-similarity for
different parameter values. Other formalizations of perceptual similarity are possible,
including ones that allow for different discriminability in different areas of the state space,
but we stick with Nosofsky's similarity function for the time being, because it is
mathematically simple, yet an established notion in mathematical psychology.

@@

As for utility functions, we take another Nosofsky-style function:
\begin{align*}
  \util(\mystate{i}, \mystate{j} \myts ; \myts \toler) =
      \begin{cases}
    1 & \textrm{if } \toler = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
    0 & \textrm{if } \toler = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
 \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^2}{ \toler^2} \right ) & \textrm{otherwise.} \\
    \end{cases}
\end{align*}
The tolerance parameter $\toler \ge 0$ models the amount of tolerable pragmatic slack or
communicative imprecision. This choice of utility function is governed partly by convenience in
parallel to the similarity function, but also because it has, as we believe, the right general
properties for a communicative payoff function. Unlike utilities that are, say, linearly or
quadratically decreasing in physical distance
\citep[c.f.][]{JagerMetzger2011:Voronoi-Languag,FrankeJager2010:Vagueness-Signa}, utilities
that are exponentially decreasing in negative quadratic distance can model situations where a
small amount of imprecision in communication is tolerable, whereas similarly small differences
in intolerably far away interpretations matter very little, with a smooth transition between
these regimes \citep[c.f.][]{OConnor2013:The-Evolution-o}.

@@

  \caption{Example strategies under \rdd at stopping time. Each line corresponds to a message and plots, for each state, the probability that the message is used.}
  \label{fig:example_strats}
\end{figure}

@@

\subsection{Evolutionary benefits of imprecision}
The inclusion of confusability of similar states has noteworthy effects on the evolving meaning of signals.
It transpired from our results that imprecision can have further accelerating and, surprisingly, unifying effects on meaning evolution.
The unifying property of perceptual imprecision could be considered an evolutionary beneficial
side-effect. A certain degree of imprecision can lead to higher \emph{within group expected utility},
defined as the average expected utility that each evolved language scored when playing against
an arbitrary other language obtained for the same parameter
values. Figure~\ref{fig:GroupMeasuresB} gives a representative example. The observation repeats
for other parameter values: while imprecision might decrease the communicative efficiency of
individual languages, it increases the conceptual coherence and communicative success between
independently evolving strategies. It is as if mere confusability of states imposes a
regularity constraint on evolving categories.

>>>>

Many concepts and expressions are vague. A vague category knows clear
cases that fall under it, clear cases that do not, and also so-called
borderline cases. Borderline cases do not clearly apply, nor do they
clearly not apply, and there may be differences between borderline
cases in terms of how well they represent the category in
question. Vagueness does not seem to dramatically affect the success
of everyday communication, but it is troublesome for some of the most
prominent theories of language and meaning. This is especially so for the logico-semantic tradition of Frege, Russell and the young
Wittgenstein which is challenged by the paradoxes vagueness gives rise
to.

@@

The next section introduces the background against which the work presented here can be
appreciated. Section~\ref{sec:repl-diff-dynam} introduces a generalization of the replicator
dynamic which is derived from the idea that agents imitate other agents' behavior while
possibly confusing similar states. This imprecise imitation dynamic is explored in
Section~\ref{sec:exploring-rdd}. Section~\ref{sec:discussion} reflects and compares our
approach to others. Appendix~\ref{sec:repl-diff-dynam} provides formal detail.

@@

@@

  \caption{A round of play in a sim-max game: Nature (N) chooses a state $\mystate{1}$ with
    probability $\Pr(\mystate{1})$; a sender (S) selects a message
    $\messg = \Spure(\mystate{1})$ determined by his pure strategy $\Spure$; a receiver (R)
    selects a state $\mystate{2} = \Rpure(\messg)$ determined by her pure strategy
    $\Rpure$. Payoff for both sender and receiver is given by
    $\Utils(\mystate{1}, \mystate{2})$.}
  \label{fig:sim-max-illustration}
\end{figure}

Formally, a sim-max game consists of a set of states $\States$, a set of messages $\Messgs$ typically with much
fewer messages than states, a probability distribution $\Pr \in \Delta(\States)$ such that
$\Pr(\state)$ gives the probability that state $\state$ occurs, a similarity metric\footnote{A
  metric is a function of distance between any two points in a given space. It should satisfy
  certain axioms that ensure behavior that one would intuitively expect from the words
  ``similarity'' and ``distance'' alone, but these details do not matter for the purposes of
  this paper.}
  % : similarity should always be non-negative, zero only for the distance
  % between an element and itself, symmetric, and satisfy the triangle inequality that
  % $\similarity(x,z) \le \similarity(x,y) + \similarity(y,z)$
on states $\similarity \mycolon \States \times \States \rightarrow \mathds{R}$ such that
$\similarity(\mystate{1}, \mystate{2})$ is the (perceptual) similarity between $\mystate{1}$
and $\mystate{2}$, and a utility function $\Utils \mycolon \States \times \States \rightarrow
\mathds{R}$ such that $\Utils(\mystate{1}, \mystate{2})$ is the payoff for sender and receiver
for a play with actual state $\mystate{1}$ and receiver interpretation $\mystate{2}$. We
identify the receiver's acts with the states of the world, so that the game is one of guessing
the actual state, so to speak. The utility function should be a monotonically decreasing
function of similarity.
In its original formulation, a sender's behavior is represented by a pure strategy $\Spure \in \Messgs^\States$ which associates states to messages (given a state $\state \in \States$ which message $\messg \in \Messgs$ would be sent), and a receiver's behavior by a pure strategy $\Rpure \in \States^\Messgs$ which does the opposite.
A play in such a game is illustrated in Figure~\ref{fig:sim-max-illustration}.

@@

Seen in this light, sim-max games may provide a foundation to approaches in cognitive semantics
that rely on the notion of conceptual spaces.
\citet[][70--77]{Gardenfors2000:Conceptual-Spac}, for example, has prominently argued that
natural categories are convex regions in conceptual space. If the conceptual space has a
suitable metric, convex categories can be derived from a set of prototypes. The category
corresponding to prototype $p$ is the set of points that are more similar to $p$ than to any
other. In this way, \citet{Gardenfors2000:Conceptual-Spac} argues, an efficient categorization
system can be obtained: storing the prototypes lets us recover the categories without having to
store each category's extension. However, what is left unexplained, is where the prototypes
come from, and why we would not see just any distribution of prototypes as an equally efficient
classification system. This is where sim-max games can contribute a principled approach to
deriving, in an independent way, not only convex categories but also prototypical exemplars
belonging to them.  These ideas, and more, are developed further by
\citet{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag,OConnor2014-OCOEPC},
among others.

@@

An alternative approach is taken by \citet{FrankeJager2010:Vagueness-Signa} and
\citet{OConnor2013:The-Evolution-o} who show how the above desiderata can be met by evolving
strategies in sim-max games. The approach we take here is similar in spirit, but different in
relevant detail. A more in-depth comparison is deferred until Section~\ref{sec:discussion}. Let
us first look into our own proposal in more detail.

@@

Signaling agents can adapt their dispositions to act, given some feedback about their past
success, in multiple ways. Usually, we would assume that changes in behavior should, at least
on average, lean towards increasing chances of communicative success. Such behavioral
adaptations can be described at different levels of abstraction. At the level of individual
agents, we can picture a more or less idealized process of how each agent adapts dispositions
for future actions based on various pieces of information available to the agent. More
abstractly, at the level of a population of agents, we can describe how average behavioral
dispositions will evolve. The population-level perspective abstracts over petty stochastic
fluctuations and zooms in on the general tendency or direction of evolution that ensues from
behavior at the agent-level.

@@

The \emph{expected utility} of choices at each choice point is:\footnote{The notation
  $\propto$, for ``proportional to'', that is used here and hereafter means that the right-hand
  side must possibly yet be normalized so as to have probabilities sum to one. In general,
  writing $P(x) \propto f(x)$ for any function $f \mycolon X \rightarrow \mathds{R}$ is
  shorthand for $P(x) = \frac{f(x)}{\sum_{x' \in X} f(x')}$. }
\begin{align*}
  \EU(\messg, \state, \Rstrat) & = \sum_{\state \in \States}
  \Rstrat(\state' \probbar \messg) \myts U(\state, \state') \\
  \EU(\state', \messg, \Sstrat) & = \sum_{\state \in
    \States} P(\state \probbar \messg)  \myts
  U(\state, \state')\,, \text{\ \  where } &
P(\state \probbar \messg) & \propto \Pr(\state) \myts \Sstrat(\messg \probbar \state) \,.
\end{align*}
Expected utilities are sums, over all possible concrete outcomes, of the utilities of these
outcomes, weighted by how likely they occur. For instance, the expected utility for the sender
of choosing message $\messg$, given state $\state$ and a receiver strategy $\Rstrat$ is the
sum, for each interpretation $\state'$, of the probability that some random receiver will
choose $\state'$ given $\messg$ times the actual utility of triple $\state$, $\messg$ and
$\state'$.

@@

The replicator dynamic is an abstract population-level dynamic that describes the mean expected
change of behavioral dispositions in a population of signalers. There are several ways of
deriving the replicator dynamic from agent-level processes of behavioral adaptation. We focus
here on one of the simplest: \emph{imitation of success}
\citep[see][]{Sandholm2010:Population-Game}. The intuitive idea is the following (see
Appendix~XYZ for a derivation of the standard replicator dynamic from this update scheme).
Every now and then a random agent gets a chance to alter his behavior for one of his choice
points. Say, this is a sender who gets to ``reconsider'' his choice of message for state
$\state$. The revising agent then observes what some random agent does at $\state$, say
$\messg$, and will henceforth play $\messg$ in $\state$ with a probability given by the
expected utility of $\messg$ for state $\state$.

@@

The presence of observation and realization noise also affects imitation of successes. Here, we
focus on the main ideas. Formal detail is provided by Appendix~XYZ. Take a sender who gets to
revise behavior for state $\state$. What agents can plausibly revise by imitation is their pure
strategy, which maps perceived states onto messages. We should therefore assume that when an
agent gets to revise his strategy for the perceived state $\state$, this need not necessarily
be the actual state. Moreover, when that agent observes $\state$, another possibly
to-be-imitated agent may yet perceive a completely different state altogether. Given what that
latter agent perceives, his actual (pure) strategy will determine what he plays. In sum, to
describe how likely the potential imitator observes a message choice $\messg$, we are
interested in the conditional probability $P_o(\messg \probbar \state)$ that some other random
agent realizes $\messg$ when the first agent perceived $\state$. This
$P_o(\messg \probbar \state)$ is derived from the prior probability of states, the current
sender population behavior $\Sstrat$ and the given observation noise (see
Appendix~XYZ). Eventually, the imitating agent adopts $\messg$ as his choice for $\state$ with
probability given by the expected utility of sending $\messg$ when perceiving state
$\state$. Expected utility should, of course, take the probabilistic confusability of states
into account as well.

@@

Appendix~XYZ shows how imprecise imitation of this sort leads to mean changes in population
frequencies of choices that can be covered by the following discrete-time formulation:
\begin{align}
  \label{eq:nRD}
  \Sstrat'(\messg \probbar \state) & \propto P_o(\messg \probbar \state) \myts \EU(\messg,
  \state, \Rstrat)\,, & \Rstrat'(\state \probbar \messg) & \propto P_o(\state \probbar \messg)
  \myts \EU(\state, \messg, \Sstrat)\,.
\end{align}
\noindent This looks very much like the discrete-time formulation of the standard replicator
dynamic in (\ref{eq:RDDiscrete}), but there are, of course, the aforementioned
differences. Firstly, expected utility here takes stochastic confusability of states into
account. Secondly, where the standard replicator dynamic had probabilities
$\Sstrat(\messg \probbar \state)$ and $\Rstrat(\state \probbar \messg)$, we now have
$P_o(\messg \probbar \state)$ and $P_o(\state \probbar \messg)$ respectively. If there is no
observation or realization noise, $P_o(\messg \probbar \state)$ reduces to
$\sigma(\messg \probbar \state)$ and $P_o(\state \probbar \messg)$ reduces to
$\Rstrat(\state \probbar \messg)$. The noisy replicator dynamic in (\ref{eq:nRD})
conservatively extends the classic case in (\ref{eq:RDDiscrete}).

@@

Let the state space consist of $n \ge 2$ states that are equally spaced across the unit
interval, including $0$ and $1$. All states occur, for simplicity, with the same probability.
The distance $\card{\mystate{i} - \mystate{j}}$ is the objective, physical similarity between
two states $\mystate{i}$ and $\mystate{j}$. Distance in physical space feeds into a perceptual
similarity function, as described by \citet{Nosofsky1986:Attention-Simil}:
\begin{align*}
  \similarity(\mystate{i}, \mystate{j} \myts ; \myts \imprecision) =
      \begin{cases}
    1 & \textrm{if } \imprecision = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
    0 & \textrm{if } \imprecision = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
 \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \imprecision^2} \right ) & \textrm{otherwise,} \\
    \end{cases}
\end{align*}
where $\imprecision \ge 0$ is an imprecision parameter. When $\imprecision=0$ agents perfectly
discriminate between states; when $\imprecision \rightarrow \infty$ agents cannot discriminate
states at all. Figure~\ref{fig:NosofskySim} gives an impression of Nosofsky-similarity for
different parameter values. Other formalizations of perceptual similarity are possible,
including ones that allow for different discriminability in different areas of the state space,
but we stick with Nosofsky's similarity function for the time being, because it is
mathematically simple, yet an established notion in mathematical psychology.

@@

As for utility functions, we take another Nosofsky-style function:
\begin{align*}
  \util(\mystate{i}, \mystate{j} \myts ; \myts \toler) =
      \begin{cases}
    1 & \textrm{if } \toler = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
    0 & \textrm{if } \toler = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
 \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \toler^2} \right ) & \textrm{otherwise.} \\
    \end{cases}
\end{align*}
The tolerance parameter $\toler \ge 0$ models the amount of tolerable pragmatic slack or
communicative imprecision. This choice of utility function is governed partly by convenience in
parallel to the similarity function, but also because it has, as we believe, the right general
properties for a communicative payoff function. Unlike utilities that are, say, linearly or
quadratically decreasing in physical distance
\citep[c.f.][]{JagerMetzger2011:Voronoi-Languag,FrankeJager2010:Vagueness-Signa}, utilities
that are exponentially decreasing in negative quadratic distance can model situations where a
small amount of imprecision in communication is tolerable, whereas similarly small differences
in intolerably far away interpretations matter very little, with a smooth transition between
these regimes \citep[c.f.][]{OConnor2013:The-Evolution-o}.

@@

  \caption{Example strategies under \rdd at stopping time of our simulations.}
  \label{fig:example_strats}
\end{figure}

@@

\subsection{Evolutionary benefits of imprecision}
The inclusion of confusability of similar states has noteworthy effects on the evolving meaning of signals.
It transpired from our results that imprecision can have further accelerating and, surprisingly, unifying effects on meaning evolution.
The unifying property of perceptual imprecision could be considered an evolutionary beneficial
side-effect. Higher imprecision can lead to higher \emph{within group expected utility},
defined as the average expected utility that each evolved language scored when playing against
an arbitrary other language obtained for the same parameter
values. Figure~\ref{fig:GroupMeasuresB} gives a representative example. The observation repeats
for other parameter values: while imprecision might decrease the communicative efficiency of
individual languages, it increases the conceptual coherence and communicative success between
independently evolving strategies. It is as if mere confusability of states imposes a
regularity constraint on evolving categories.
